---
title: "Lab-06-Dodson"
author: "Kenaniah Dodson"
date: "2025-05-05"
output: 
  github_document:
    toc: true
    toc_depth: 3
    preserve_yaml: true
  rmdformats::downcute:
    toc_depth: 3
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: "tango"
    code_folding: show
always_allow_html: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, eval=FALSE, echo=FALSE}
# Library
library(report)    # package for statistical report writing and citations
library(tidyverse) # data wrangling
# Load packages
library(renv)         # R environment management
library(devtools)     # R developer tools
library(here)         # relative filepaths for reproducibility
library(tidyverse)    # data wrangling
library(stringi)      # string wrangling
library(kableExtra)   # table formatting
library(tidycensus)   # census data
library(ggplot2)      # data visualization
library(patchwork)    # data visualization layout
library(magick)       # image editing in R
library(ggiraph)      # interactive data visualization
library(htmlwidgets)  # create widget from interactive visualizations
library(htmltools)    # encode HTML code in map tooltips
library(widgetframe)  # save widget of interactive visualizations
library(showtext)     # utilize special fonts in graphs
library(scales)       # format numbers and colors for graphs
library(tigris)       # pull US Census shapefiles
library(cowplot)      # data visualization plotting
library(gridExtra)    # grid for data visualizations
library(biscale)      # bivariate mapping
library(cluster)      # clustering algorithms
library(factoextra)   # clustering algorithms & visualization
library(rio)          # read excel file from URL
library(unhcrthemes)  # data visualization themes
library(ggrepel)      # data visualization formatting to avoid overlapping
library(rcompanion)   # data visualization of variable distribution
library(ggpubr)       # data visualization of variable distribution
library(moments)      # measures of skewness and kurtosis
library(tinytable)    # format regression tables
library(modelsummary) # format regression tables
library(report)    # package for statistical report writing and citations
```


```{r, eval=FALSE}
import::here( "fips_census_regions",
              "load_svi_data",
              "merge_svi_data",
              "census_division",
              "write_model_outcomes",
             # notice the use of here::here() that points to the .R file
             # where all these R objects are created
             .from = here::here("analysis/project_data_steps_dodson.R"),
             .character_only = TRUE)
```


```{r, eval=FALSE}
# Load SVI data sets

## Raw data
svi_2010 <- readRDS(here::here("data/raw/Census_Data_SVI/svi_2010_trt10.rds"))
svi_2020 <- readRDS(here::here("data/raw/Census_Data_SVI/svi_2020_trt10.rds"))

### Divisional data
svi_2010_divisional <- load_svi_data(svi_2010, rank_by = "divisional", location = census_division, percentile=.75)
svi_2020_divisional <- load_svi_data(svi_2020, rank_by = "divisional", location =  census_division, percentile=.75)
svi_divisional <- merge_svi_data(svi_2010_divisional, svi_2020_divisional)

### National data
svi_2010_national <- load_svi_data(svi_2010, percentile=.75)
svi_2020_national <- load_svi_data(svi_2020, percentile=.75)
svi_national <- merge_svi_data(svi_2010_national, svi_2020_national)


### NMTC Data
svi_divisional_nmtc_df <- readRDS(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_divisional_nmtc.rds")))

svi_national_nmtc_df <- readRDS(file = here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_national_nmtc.rds")))

### LIHTC Data
svi_divisional_lihtc_df <- readRDS(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_divisional_lihtc.rds")))

svi_national_lihtc_df <- readRDS(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_national_lihtc.rds")))

# Load model data

load(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_did_models_nmtc.RData")))

load(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_did_models_lihtc.RData")))
```


# Introduction

This project investigates the relationship between federal investment programs and social vulnerability in the South Atlantic Division of the United States. Specifically, it evaluates the impact of the New Markets Tax Credit (NMTC) and Low-Income Housing Tax Credit (LIHTC) programs on vulnerable census tracts between 2010 and 2020. Social vulnerability is assessed using the CDC’s Social Vulnerability Index (SVI), which includes four dimensions: socioeconomic status, household characteristics, racial and ethnic minority status, and housing and transportation access. Data were sourced from the U.S. Census Bureau, the Federal Housing Finance Agency, and were normalized to 2010 census tracts using the NHGIS crosswalk.

The project employs a mix of spatial and statistical methods, including choropleth and bivariate mapping to visualize geographic trends and relationships, correlation analyses to explore variable associations, K-means clustering to identify patterns in funding and vulnerability, and difference-in-differences (diff-in-diff) regression models to measure the effects of NMTC and LIHTC programs over time. Overall, we found no statistically significant impact of either program on social vulnerability outcomes. There was also no consistent correlation between the amount of funding received and the vulnerability of the census tracts, raising concerns about the allocation strategy. These findings suggest a need for more targeted and transparent investment practices. Future studies should investigate how and where funds were spent, and whether other non-financial factors—such as local governance or administrative capacity—played a role in determining outcomes. Stakeholders would benefit from more granular, program-level data to assess effectiveness and improve equity in resource distribution.

# Data

Data from this report is collected from the U.S. Census Bureau and the Federal Housing Finance Agency. However, since census tracts have changed between 2010 and 2020, the data from 2020 was crosswalked to 2010 tracts with the NHGIS crosswalk.
Throughout the South Atlantic Division section of this report, we will track indicators of social vulnerability using data from the U.S. Census Bureau and the Federal Housing Finance Agency. This data will be used in line withe Centers for Diseas Control (CDC)'s Social Vulnerability Index, which divides Social Vulnerability into four categories: socioeconomic status, household characterisitcs, racial and ethnic minority status, and housing type and transportation.

The SVI data sets contains 73057 tracts nationally. Divisionally, the South Atlantic Division contains 13706 tracts.
In the South Atlantic Division, the most vulnerable state was Florida both in 2010 and in 2020.
The most vulnerable by themes 1-4 in 2020 were: Broward County, Miami-Dade County, Alachua County, and Hillsborough County. 
For contrast, the most vulnerable by themes 1-4 in 2010 were: Broward County, Hillsborough County, Alachua County, and Miami-Dade County.

Tracts were identified as being eligible based on a number of factors. For the NMTC, census tract poverty rate would need to be equal or greater than 20%, Median Family Income would need to be less than or equal to 80% of the average, and tract unemployment to national unemployment ratio would need to be greater than 1.5. Tracts qualifying for NMTC awards would create a "flag". Counting up these flags allow us to see which areas were the most socially vulnerable. For the NMTC, the largest ammount of flags counted was 9936. For the LIHTC, the maximum was 2457. 

Data was grouped and observed in a number of ways, including clustering and diff-in-diff models. These models are used to show how different tracts are affected, as well as how the flag count relates to the amount of money invested into different tracts.

# Analysis

Throughout this project, we have used various different methodologies in order to best understand our data.

Spatial analysis was used to map our division into states, counties, and tracts, in order to understand how geography may have impacted the data. Choropleth maps are shown to explain population, as well as our social vulnrability flags. Bivariate mapping is used to show relationships between SVI variables. 

K-Means Clustering was used in order to discover relationships between flag counts and the amount of money invested into at-risk areas. Since there was a large variance between our maximum and minimum flag counts as well as our maximum and minimum money invested, outliers were affecting our data. Clustering allowed us to view the data in smaller groups that were similar in flag counts, and money invested.
Diff-In-Diff regression was used in various models to view how the NMTC and LIHTC programs affected our tracts between 2010 and 2020.

# Results

## NMTC Diff-In-Diff Models

### Socioeconomic SVI

```{r, eval=FALSE}
write_model_outcomes(m1_nmtc_div, "NMTC", "socioeconomic status")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_SES with treat, post and cbsa (formula: SVI_FLAG_COUNT_SES ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.17, F(131, 10358) = 16.00, p < .001, adj. R2 = 0.16)

The effect of treat × post is statistically non-significant and negative (beta = -0.15, 95% CI [-0.38, 0.08], t(10358) = -1.28, p = 0.200; Std. beta = -0.01, 95% CI [-0.03, 6.08e-03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on socioeconomic status-related social vulnerability and economic outcomes.

### Household Characteristics SVI

```{r, eval=FALSE}
write_model_outcomes(m2_nmtc_div, "NMTC", "household characteristics")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_HHCHAR with treat, post and cbsa (formula: SVI_FLAG_COUNT_HHCHAR ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and weak proportion of variance (R2 = 0.10, F(131, 10358) = 8.60, p < .001, adj. R2 = 0.09)

The effect of treat × post is statistically non-significant and negative (beta = -0.13, 95% CI [-0.29, 0.03], t(10358) = -1.59, p = 0.111; Std. beta = -0.01, 95% CI [-0.03, 3.44e-03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on household characteristics-related social vulnerability and economic outcomes.

### Racial and Ethnic Minority SVI

```{r, eval=FALSE}
write_model_outcomes(m3_nmtc_div, "NMTC", "racial and ethnic minority status")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_REM with treat, post and cbsa (formula: SVI_FLAG_COUNT_REM ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.25, F(131, 10358) = 25.92, p < .001, adj. R2 = 0.24)

The effect of treat × post is statistically non-significant and negative (beta = -0.02, 95% CI [-0.09, 0.06], t(10358) = -0.43, p = 0.667; Std. beta = -3.67e-03, 95% CI [-0.02, 0.01])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on racial and ethnic minority status-related social vulnerability and economic outcomes.

### Housing and Transportation SVI

```{r, eval=FALSE}
write_model_outcomes(m4_nmtc_div, "NMTC", "housing and transportation access")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_HOUSETRANSPT with treat, post and cbsa (formula: SVI_FLAG_COUNT_HOUSETRANSPT ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and weak proportion of variance (R2 = 0.08, F(131, 10358) = 6.93, p < .001, adj. R2 = 0.07)

The effect of treat × post is statistically non-significant and positive (beta = 4.97e-03, 95% CI [-0.16, 0.17], t(10358) = 0.06, p = 0.953; Std. beta = 5.51e-04, 95% CI [-0.02, 0.02])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on housing and transportation access-related social vulnerability and economic outcomes.

### Overall SVI

```{r, eval=FALSE}
write_model_outcomes(m5_nmtc_div, "NMTC", "socioeconomic, household characteristics, racial and ethnic minority status, and housing and transportation access")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_OVERALL with treat, post and cbsa (formula: SVI_FLAG_COUNT_OVERALL ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.18, F(131, 10358) = 16.97, p < .001, adj. R2 = 0.17)

The effect of treat × post is statistically non-significant and negative (beta = -0.29, 95% CI [-0.75, 0.16], t(10358) = -1.26, p = 0.207; Std. beta = -0.01, 95% CI [-0.03, 6.23e-03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on socioeconomic, household characteristics, racial and ethnic minority status, and housing and transportation access-related social vulnerability and economic outcomes.

### Median Income Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m6_nmtc_div, "NMTC", "Median Income")
```
We fitted a linear model (estimated using OLS) to predict MEDIAN_INCOME with treat, post and cbsa (formula: MEDIAN_INCOME ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.20, F(131, 10350) = 19.93, p < .001, adj. R2 = 0.19)

The effect of treat × post is statistically significant and positive (beta = 0.06, 95% CI [0.01, 0.11], t(10350) = 2.54, p = 0.011; Std. beta = 0.02, 95% CI [5.08e-03, 0.04])

Since the effect of treat x post is statistically significant, we can conclude that the NMTC program had a measurable impact on Median Income-related social vulnerability and economic outcomes.

### Median Home Value Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m7_nmtc_div, "NMTC", "Median Home Value")
```
We fitted a linear model (estimated using OLS) to predict MEDIAN_HOME_VALUE with treat, post and cbsa (formula: MEDIAN_HOME_VALUE ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.42, F(131, 10116) = 56.78, p < .001, adj. R2 = 0.42)

The effect of treat × post is statistically non-significant and positive (beta = 0.02, 95% CI [-0.04, 0.09], t(10116) = 0.67, p = 0.503; Std. beta = 5.06e-03, 95% CI [-9.74e-03, 0.02])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on Median Home Value-related social vulnerability and economic outcomes.

### House Price Index Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m8_nmtc_div, "NMTC", "House Price Index")
```
We fitted a linear model (estimated using OLS) to predict HOUSE_PRICE_INDEX with treat, post and cbsa (formula: HOUSE_PRICE_INDEX ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.37, F(123, 4962) = 23.82, p < .001, adj. R2 = 0.36)

The effect of treat × post is statistically non-significant and positive (beta = 0.05, 95% CI [-0.04, 0.13], t(4962) = 1.05, p = 0.292; Std. beta = 0.01, 95% CI [-0.01, 0.03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on House Price Index-related social vulnerability and economic outcomes.

## LIHTC Diff-In-Diff Models

### Socioeconomic SVI

```{r, eval=FALSE}
write_model_outcomes(m1_lihtc_div, "LIHTC", "socioeconomic status")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_SES with treat, post and cbsa (formula: SVI_FLAG_COUNT_SES ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.26, F(77, 928) = 4.33, p < .001, adj. R2 = 0.20)

The effect of treat × post is statistically non-significant and positive (beta = 0.03, 95% CI [-0.37, 0.42], t(928) = 0.14, p = 0.888; Std. beta = 3.98e-03, 95% CI [-0.05, 0.06])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on socioeconomic status-related social vulnerability and economic outcomes.

### Household Characteristics SVI

```{r, eval=FALSE}
write_model_outcomes(m2_lihtc_div, "LIHTC", "household characteristics")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_HHCHAR with treat, post and cbsa (formula: SVI_FLAG_COUNT_HHCHAR ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.24, F(77, 928) = 3.86, p < .001, adj. R2 = 0.18)

The effect of treat × post is statistically non-significant and negative (beta = -0.12, 95% CI [-0.42, 0.18], t(928) = -0.77, p = 0.442; Std. beta = -0.02, 95% CI [-0.08, 0.03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on household characteristics-related social vulnerability and economic outcomes.

### Racial and Ethnic Minority SVI

```{r, eval=FALSE}
write_model_outcomes(m3_lihtc_div, "LIHTC", "racial and ethnic minority status")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_REM with treat, post and cbsa (formula: SVI_FLAG_COUNT_REM ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.37, F(77, 928) = 6.99, p < .001, adj. R2 = 0.31)

The effect of treat × post is statistically non-significant and positive (beta = 2.18e-03, 95% CI [-0.12, 0.12], t(928) = 0.04, p = 0.972; Std. beta = 9.29e-04, 95% CI [-0.05, 0.05])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on racial and ethnic minority status-related social vulnerability and economic outcomes.

### Housing and Transportation SVI

```{r, eval=FALSE}
write_model_outcomes(m4_lihtc_div, "LIHTC", "housing and transportation access")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_HOUSETRANSPT with treat, post and cbsa (formula: SVI_FLAG_COUNT_HOUSETRANSPT ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.17, F(77, 928) = 2.44, p < .001, adj. R2 = 0.10)

The effect of treat × post is statistically non-significant and negative (beta = -0.02, 95% CI [-0.30, 0.27], t(928) = -0.12, p = 0.903; Std. beta = -3.65e-03, 95% CI [-0.06, 0.06])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on housing and transportation access-related social vulnerability and economic outcomes.

### Overall SVI

```{r, eval=FALSE}
write_model_outcomes(m5_lihtc_div, "LIHTC", "socioeconomic, household characteristics, racial and ethnic minority status, and housing and transportation access")
```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_OVERALL with treat, post and cbsa (formula: SVI_FLAG_COUNT_OVERALL ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.28, F(77, 928) = 4.65, p < .001, adj. R2 = 0.22)

The effect of treat × post is statistically non-significant and negative (beta = -0.10, 95% CI [-0.83, 0.62], t(928) = -0.28, p = 0.777; Std. beta = -7.92e-03, 95% CI [-0.06, 0.05])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on socioeconomic, household characteristics, racial and ethnic minority status, and housing and transportation access-related social vulnerability and economic outcomes.

### Median Income Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m6_lihtc_div, "LIHTC", "Median Income")
```
We fitted a linear model (estimated using OLS) to predict MEDIAN_INCOME with treat, post and cbsa (formula: MEDIAN_INCOME ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.31, F(77, 926) = 5.47, p < .001, adj. R2 = 0.26)

The effect of treat × post is statistically non-significant and negative (beta = -4.13e-03, 95% CI [-0.13, 0.12], t(926) = -0.06, p = 0.949; Std. beta = -1.73e-03, 95% CI [-0.06, 0.05])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on Median Income-related social vulnerability and economic outcomes.

### Median Home Value Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m7_lihtc_div, "LIHTC", "Median Home Value")
```
We fitted a linear model (estimated using OLS) to predict MEDIAN_HOME_VALUE with treat, post and cbsa (formula: MEDIAN_HOME_VALUE ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.41, F(77, 880) = 8.00, p < .001, adj. R2 = 0.36)

The effect of treat × post is statistically non-significant and positive (beta = 4.72e-03, 95% CI [-0.17, 0.18], t(880) = 0.05, p = 0.958; Std. beta = 1.36e-03, 95% CI [-0.05, 0.05])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on Median Home Value-related social vulnerability and economic outcomes.

### House Price Index Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m8_lihtc_div, "LIHTC", "House Price Index")
```
We fitted a linear model (estimated using OLS) to predict HOUSE_PRICE_INDEX with treat, post and cbsa (formula: HOUSE_PRICE_INDEX ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.40, F(32, 241) = 4.98, p < .001, adj. R2 = 0.32)

The effect of treat × post is statistically non-significant and positive (beta = 0.09, 95% CI [-0.19, 0.36], t(241) = 0.64, p = 0.523; Std. beta = 0.03, 95% CI [-0.07, 0.13])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on House Price Index-related social vulnerability and economic outcomes.

# Discussion and Recommendations

From our bivariate maps, we can see that there is a high ratio of flags to population in West Virginia, eastern North Carolina, southern Georgia, and across Florida. We can also see that the tracts appear to be very similar between 2010 and 2020. Although this model does not show total populations and may not be the best way to view this data, it could suggest that the programs did not have the expected affect.

Suprisingly, when we look at our amount of flags compared to the amount of money invested, we see that there is not a strong correlation. This would suggest that the amount of money being spent may not have been decided based on social vulnerability factors soley. Due to this bias in our data, we viewed our data in groups so that we may see how different flags and spending levels correlate. However, even in this model, we cannot conclude that the amount of money spent had a significant impact on social vulnerability due to the amount of flags between 2010 and 2020 remaining similar.

Finally, when we viewed the impact of both the LIHTC and NMTC programs onto different social vulnerability outcomes, we cannot conclude that either programs had an impact on any of the outcomes.

Due to the large varience in flags and amount of money spent, as well as the lack of significance in the outcomes, it would be beneficial to study further into how the money was spent specifically. It does not appear as if funding was going where it was most needed. Although this study could not conclude that social vulnerability flags were related to the amount of money invested, it would be interesting to see what other factors or decision makers were impactful. Understanding how the money was spent could unlock further understanding into how the money could be used to see a better result.

# References

## Packages

```{r, eval=FALSE}
session <- sessionInfo()

report_system(session)
```
Analyses were conducted using the R Statistical language (version 4.4.0; R Core Team, 2024) on Windows 10 x64 (build 19045)

```{r, eval=FALSE}
cite_packages(session, include_R = TRUE)
```
- Arel-Bundock V (2022). “modelsummary: Data and Model Summaries in R.” _Journal of Statistical Software_, *103*(1), 1-23. doi:10.18637/jss.v103.i01 <https://doi.org/10.18637/jss.v103.i01>.
  - Arel-Bundock V (2025). _tinytable: Simple and Configurable Tables in 'HTML', 'LaTeX', 'Markdown', 'Word', 'PNG', 'PDF', and 'Typst' Formats_. R package version 0.8.0, <https://CRAN.R-project.org/package=tinytable>.
  - Auguie B (2017). _gridExtra: Miscellaneous Functions for "Grid" Graphics_. R package version 2.3, <https://CRAN.R-project.org/package=gridExtra>.
  - Chan C, Leeper T, Becker J, Schoch D (2023). _rio: A Swiss-army knife for data file I/O_. <https://cran.r-project.org/package=rio>.
  - Cheng J, Sievert C, Schloerke B, Chang W, Xie Y, Allen J (2024). _htmltools: Tools for HTML_. R package version 0.5.8.1, <https://CRAN.R-project.org/package=htmltools>.
  - Gagolewski M (2022). “stringi: Fast and portable character string processing in R.” _Journal of Statistical Software_, *103*(2), 1-59. doi:10.18637/jss.v103.i02 <https://doi.org/10.18637/jss.v103.i02>.
  - Gohel D, Skintzos P (2025). _ggiraph: Make 'ggplot2' Graphics Interactive_. R package version 0.8.13, <https://CRAN.R-project.org/package=ggiraph>.
  - Grolemund G, Wickham H (2011). “Dates and Times Made Easy with lubridate.” _Journal of Statistical Software_, *40*(3), 1-25. <https://www.jstatsoft.org/v40/i03/>.
  - Karambelkar B (2017). _widgetframe: 'Htmlwidgets' in Responsive 'iframes'_. R package version 0.3.1, <https://CRAN.R-project.org/package=widgetframe>.
  - Kassambara A (2023). _ggpubr: 'ggplot2' Based Publication Ready Plots_. R package version 0.6.0, <https://CRAN.R-project.org/package=ggpubr>.
  - Kassambara A, Mundt F (2020). _factoextra: Extract and Visualize the Results of Multivariate Data Analyses_. R package version 1.0.7, <https://CRAN.R-project.org/package=factoextra>.
  - Komsta L, Novomestky F (2022). _moments: Moments, Cumulants, Skewness, Kurtosis and Related Tests_. R package version 0.14.1, <https://CRAN.R-project.org/package=moments>.
  - Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2025). _cluster: Cluster Analysis Basics and Extensions_. R package version 2.1.8.1 - For new features, see the 'NEWS' and the 'Changelog' file in the package source), <https://CRAN.R-project.org/package=cluster>.
  - Makowski D, Lüdecke D, Patil I, Thériault R, Ben-Shachar M, Wiernik B (2023). “Automated Results Reporting as a Practical Tool to Improve Reproducibility and Methodological Best Practices Adoption.” _CRAN_. <https://easystats.github.io/report/>.
  - Mangiafico SS (2025). _rcompanion: Functions to Support Extension Education Program Evaluation_. Rutgers Cooperative Extension, New Brunswick, New Jersey. version 2.5.0, <https://CRAN.R-project.org/package=rcompanion/>.
  - Müller K (2020). _here: A Simpler Way to Find Your Files_. R package version 1.0.1, <https://CRAN.R-project.org/package=here>.
  - Müller K, Wickham H (2023). _tibble: Simple Data Frames_. R package version 3.2.1, <https://CRAN.R-project.org/package=tibble>.
  - Ooms J (2025). _magick: Advanced Graphics and Image-Processing in R_. R package version 2.8.6, <https://CRAN.R-project.org/package=magick>.
  - Pedersen T (2024). _patchwork: The Composer of Plots_. R package version 1.3.0, <https://CRAN.R-project.org/package=patchwork>.
  - Prener C, Grossenbacher T, Zehr A (2022). _biscale: Tools and Palettes for Bivariate Thematic Mapping_. R package version 1.0.0, <https://CRAN.R-project.org/package=biscale>.
  - Qiu Y, details. aotifSfAf (2020). _showtextdb: Font Files for the 'showtext' Package_. R package version 3.0, <https://CRAN.R-project.org/package=showtextdb>.
  - Qiu Y, details. aotifSfAf (2024). _sysfonts: Loading Fonts into R_. R package version 0.8.9, <https://CRAN.R-project.org/package=sysfonts>.
  - Qiu Y, details. aotisSfAf (2024). _showtext: Using Fonts More Easily in R Graphs_. R package version 0.9-7, <https://CRAN.R-project.org/package=showtext>.
  - R Core Team (2024). _R: A Language and Environment for Statistical Computing_. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.
  - Slowikowski K (2024). _ggrepel: Automatically Position Non-Overlapping Text Labels with 'ggplot2'_. R package version 0.9.6, <https://CRAN.R-project.org/package=ggrepel>.
  - Ushey K, Wickham H (2025). _renv: Project Environments_. R package version 1.1.4, <https://CRAN.R-project.org/package=renv>.
  - Vaidyanathan R, Xie Y, Allaire J, Cheng J, Sievert C, Russell K (2023). _htmlwidgets: HTML Widgets for R_. R package version 1.6.4, <https://CRAN.R-project.org/package=htmlwidgets>.
  - Vidonne C, Dicko A (2025). _unhcrthemes: UNHCR 'ggplot2' Theme and Colour Palettes_. R package version 0.6.3, <https://CRAN.R-project.org/package=unhcrthemes>.
  - Walker K (2025). _tigris: Load Census TIGER/Line Shapefiles_. R package version 2.2.0, commit 99b45062a24facbdbfdb6749ca531811d64a52b6, <https://github.com/walkerke/tigris>.
  - Walker K, Herman M (2024). _tidycensus: Load US Census Boundary and Attribute Data as 'tidyverse' and 'sf'-Ready Data Frames_. R package version 1.6.7, <https://CRAN.R-project.org/package=tidycensus>.
  - Wickham H (2016). _ggplot2: Elegant Graphics for Data Analysis_. Springer-Verlag New York. ISBN 978-3-319-24277-4, <https://ggplot2.tidyverse.org>.
  - Wickham H (2023). _forcats: Tools for Working with Categorical Variables (Factors)_. R package version 1.0.0, <https://CRAN.R-project.org/package=forcats>.
  - Wickham H (2023). _stringr: Simple, Consistent Wrappers for Common String Operations_. R package version 1.5.1, <https://CRAN.R-project.org/package=stringr>.
  - Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). “Welcome to the tidyverse.” _Journal of Open Source Software_, *4*(43), 1686. doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.
  - Wickham H, Bryan J, Barrett M, Teucher A (2024). _usethis: Automate Package and Project Setup_. R package version 3.1.0, <https://CRAN.R-project.org/package=usethis>.
  - Wickham H, François R, Henry L, Müller K, Vaughan D (2023). _dplyr: A Grammar of Data Manipulation_. R package version 1.1.4, <https://CRAN.R-project.org/package=dplyr>.
  - Wickham H, Henry L (2025). _purrr: Functional Programming Tools_. R package version 1.0.4, <https://CRAN.R-project.org/package=purrr>.
  - Wickham H, Hester J, Bryan J (2024). _readr: Read Rectangular Text Data_. R package version 2.1.5, <https://CRAN.R-project.org/package=readr>.
  - Wickham H, Hester J, Chang W, Bryan J (2022). _devtools: Tools to Make Developing R Packages Easier_. R package version 2.4.5, <https://CRAN.R-project.org/package=devtools>.
  - Wickham H, Pedersen T, Seidel D (2023). _scales: Scale Functions for Visualization_. R package version 1.3.0, <https://CRAN.R-project.org/package=scales>.
  - Wickham H, Vaughan D, Girlich M (2024). _tidyr: Tidy Messy Data_. R package version 1.3.1, <https://CRAN.R-project.org/package=tidyr>.
  - Wilke C (2024). _cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2'_. R package version 1.1.3, <https://CRAN.R-project.org/package=cowplot>.
  - Zhu H (2024). _kableExtra: Construct Complex Table with 'kable' and Pipe Syntax_. R package version 1.4.0, <https://CRAN.R-project.org/package=kableExtra>.

## Data

-  CDFI Fund (2023). _FY 2023 NMTC Public Data Release: 2003-2021 Data File Updated - Aug 21, 2023_. https://www.cdfifund.gov/documents/data-releases

-  Centers for Disease Control and Prevention/ Agency for Toxic Substances and Disease Registry/ Geospatial Research, Analysis, and Services Program. (2022). _CDC/ATSDR Social Vulnerability Index 2020 Methodology_. https://web.archive.org/web/20241028180954/https://www.atsdr.cdc.gov/placeandhealth/svi/documentation/SVI_documentation_2020.html

-  FHFA (n.d.). _HPI® Census Tracts (Developmental Index; Not Seasonally Adjusted)_. https://www.fhfa.gov/DataTools/Downloads/Pages/House-Price-Index-Datasets.aspx#atvol

-  HUD User (n.d.). _2010, 2011, and 2012 QCT data for all of the census tracts in the United States and Puerto Rico (qct_data_2010_2011_2012.xlsx)_. https://www.huduser.gov/portal/datasets/qct.html#year2010

-  HUD User (2023). _Low-Income Housing Tax Credit (LIHTC): Property Level Data_. https://www.huduser.gov/portal/datasets/lihtc/property.html

-  Novogradac New Markets Tax Credit Resource Center. (2017). _New Markets Tax Credit Low-Income Community Census Tracts - American Community Survey 2011-2015_. https://www.novoco.com/resource-centers/new-markets-tax-credits/data-tables

-  Steven Manson, Jonathan Schroeder, David Van Riper, Katherine Knowles, Tracy Kugler, Finn Roberts, and Steven Ruggles. _IPUMS National Historical Geographic Information System: Version 18.0 [2020 → 2010 Block Groups → Census Tracts Crosswalks National File]_. Minneapolis, MN: IPUMS. 2023. http://doi.org/10.18128/D050.V18.0

-  U.S. Bureau of Labor Statistics (n.d.). _CPI Inflation Calculator_. https://data.bls.gov/cgi-bin/cpicalc.pl

-  U.S. Bureau of Labor Statistics (n.d.). _QCEW County-MSA-CSA Crosswalk (For NAICS-Based Data)_. https://www.bls.gov/cew/classifications/areas/county-msa-csa-crosswalk.htm

-  U.S. Census Bureau. (2011). _2006-2010 American Community Survey 5-year_. https://www.census.gov/newsroom/releases/archives/american_community_survey_acs/cb11-208.html

-  U.S. Census Bureau. (2013). _2008-2012 American Community Survey 5-year_.  https://www.census.gov/newsroom/press-kits/2013/20131217_acs_5yr.html

-  U.S. Census Bureau. (2022). _2016-2020 American Community Survey 5-year_. https://www.census.gov/newsroom/press-releases/2022/acs-5-year-estimates.html










