---
title: "Results and Conclusions"
author: "Michelle Knopp"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    preserve_yaml: true
  rmdformats::downcute:
    toc_depth: 3
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: "tango"
    code_folding: show
always_allow_html: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=FALSE}
library(report)    # package for statistical report writing and citations
library(tidyverse) # data wrangling
```


```{r, eval=FALSE}
import::here( "fips_census_regions",
              "load_svi_data",
              "merge_svi_data",
              "census_division",
              "write_model_outcomes",
             # notice the use of here::here() that points to the .R file
             # where all these R objects are created
             .from = here::here("analysis/project_data_steps_knopp.R"),
             .character_only = TRUE)

```

```{r, eval=FALSE}
# Load SVI data sets

## Raw data
svi_2010 <- readRDS(here::here("data/raw/Census_Data_SVI/svi_2010_trt10.rds"))
svi_2020 <- readRDS(here::here("data/raw/Census_Data_SVI/svi_2020_trt10.rds"))

### Divisional data
svi_2010_divisional <- load_svi_data(svi_2010, rank_by = "divisional", location = census_division, percentile=.75)
svi_2020_divisional <- load_svi_data(svi_2020, rank_by = "divisional", location =  census_division, percentile=.75)
svi_divisional <- merge_svi_data(svi_2010_divisional, svi_2020_divisional)

### National data
svi_2010_national <- load_svi_data(svi_2010, percentile=.75)
svi_2020_national <- load_svi_data(svi_2020, percentile=.75)
svi_national <- merge_svi_data(svi_2010_national, svi_2020_national)


### NMTC Data
svi_divisional_nmtc_df <- readRDS(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_divisional_nmtc.rds")))

svi_national_nmtc_df <- readRDS(file = here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_national_nmtc.rds")))

### LIHTC Data
svi_divisional_lihtc_df <- readRDS(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_divisional_lihtc.rds")))

svi_national_lihtc_df <- readRDS(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_national_lihtc.rds")))

# Load model data

load(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_did_models_nmtc.RData")))

load(here::here(paste0("data/rodeo/", str_replace_all(census_division, " ", "_"), "_svi_did_models_lihtc.RData")))
```

# Introduction

For this project, we examined the impact of two federally funded tax programs, the New Markets Tax Program (NMTC) and the Low Income Housing Tax Credit program (LIHTC) on qualifying neighborhoods. In order to evaluate the change in neighborhoods,we will be looked at Social Variability Index (SVI) Census variables which are defined by U.S. Centers for Disease Control and Prevention(CDC)'s to measure neighborhood vulnerability. In addition, we will also looked at economic outcomes variables to include: median home values and median income from Census data, and the Federal Housing Finance Agency's house price index.

We included 2010 and 2020 data from the census bureau, utilized funding and eligibility data for the NMTC and LITC, and included data from the Federal Housing Finance Agency for the housing price index. The data was wrangled to include eligible tracts for the Mountain Division.

To analyze our data, we used both visual and statistical tools to include; waffle charts, chloropleth maps, bivariate mapping, k-means clustering, Pearson's R correlation, and diff-in-diff regression.

# Data

In order to flag social vulnerabilities for our census tracts within the Mountain Division, we pulled US Census Data. The US Centers for Disease Control and Prevention (CDC) defined appropriate variables to measure social vulnerability indices (SVI). The CDC's SVI measures how vulnerable a neighborhood is in the event of an emergency situation, such as a natural disaster or health outbreak. These variables are organized into four categories, and include the following variables:

  - Socioeconomic Status: percent living below 150% poverty, percent unemployed, population housing cost-burned, percent adults without high school diploma, percent without health insurance.
  - Household Characteristics: percent age 17 and under, percent age 65 and over, percent disabled civilians, percent single parent families, percent limited English speakers. 
  -Racial & Ethnic Minority Status: percent minority race/ethnicity.
  -Housing Type & Transportation: percent in multi-unit housing, percent in mobile housing, percent in crowded living spaces, percent with no vehicle access, percent living in group quarters.
  
We obtained data for all these categories from the US Census API for 2010 and 2020, and only used variables that were available in both decades. To account for geographic changes between the decades, we pulled census tract data from 2010 and then crosswalked it with the census block group data from 2020.

Our National data included 73,057 tracts for 2010 and 2020. The Mountain Division had a 2020 population of 24, 534, 951 living in 5,250 tracts. In 2010, the tracts with the most vulnerable (Total SVI) flags were in order, Laramie County, WY, Utah County, UT, Tooele County, UT, and 2 tracts in Salt Lake County, UT. In 2020, the most vulnerable tracts (SVI flags) included the same tracts, and did not change. The least vulnerable tracts in the Mountain Division experienced change. In 2010 the least vulnerable tracts were Summit County, UT, Glacier County, MT, and Douglas County, CO in 2010. They switched completely in 2020 to Weld County, CO, Clark County, NV, and Salt Lake County, UT. It was worth noting that Salt Lake County included tracts that were the most vulnerable and tracts that were the least vulnerable in the Mountain Division.

Next, we identified tracts that were eligible for our two tax programs; the New Markets Tax Credit (NMTC), and the Low Income Housing Tax Credit (LIHTC). Markets Tax Credits are awarded to community development entities for the purpose of investing in low income communities and recipients must meet strict criteria to be eligible, but the credits are intended to for areas with low median income and high poverty rates. Low Income Housing Tax Credits are awarded to investors with the purpose of investing in affordable housing for renters. Again, this program is designed to improve neighborhood with low gross incomes and high poverty rates. We wrangled our data to determine tracts that were deemed eligible for these tax credits, but excluded tracts that received funding prior to 2010, to insure that we are measuring the impact of the programs.

For the Mountain Division, there were 2075 tracts eligible for the NMTC program, with 56 counties receiving funding. There is a large spread within tracts receiving NMTC in the Mountain division with SVI flag counts ranging from 5 - 2558, and the NMTC dollars ranging from $800,000 to $225,215,967. Montana, Idaho, and Wyoming received high dollars compared to low flag counts. Nevada received the most dollars in the highest need counties.

For the Mountain Division, there were 57 tracts eligible for the LIHTC program, with 14 counties receiving funding. The Mountain division counties receiving LIHTC project funding between 2011-2020, had a range of 9 to 487 SVI flags in 2010. The funding ranged from $270,127 and $15,437,500. There is a large range between counties in the number of SVI flags in 2010 and funding dollars spent between 2011 and 2020. Arizona had the tracts with the highest need and the highest amount of LIHTC funding. Again, Montana and Colorado received high funding dollars for a low amount of need.

In addition to the Social Vulnerability variables, we also included 3 additional economic variables: house price index (HPI), median home value, and median income. HPI is determined by analyzing mortgage transactions. The HPI was pulled from the Federal Housing Finance Agency API. The median home value and median incomes were collected using census data.

# Analysis

We used several tools to analyze our data, both visually and statistically.

First, we used waffle charts, pictorial infographics, to look at our SVI variable categories. For each category we used 100 icons to visually show proportions of the percent of population who were vulnerable for each SVI variable, and included overall SVI for each category of variable. 

We also looked at Chloropleth maps, to look at SVI Flags to Population Ratios. In order to accomplish this, shapefiles were downloaded and SVI flags were mapped to their respective geographical county location. We included interactive tooltips so the user can get SVI information by hovering over a county.The Chloropleth maps were colored coded to visually see areas of least vulnerability (lighter color) and high vulnerablility (darker color).

BiVariate Mapping was also another visual tool to analyze the data. For our bivariate maps, we included SVI flags and the amount of funding each county received. This provided a visualization of the areas of high versus low need, and those areas who received little or high funding from our tax programs.

We also examined the correlations between funding received (tax credits) and need (SVI flags), calculating the Pearson's R. correlation coefficient. A Pearson’s R calculation of 0.62 suggests that there is a strong positive association between the dollar amount spent in NMTC from 2011-2020 and the flag count within the Mountain Division counties in 2010. Overall, within all the counties in the Mountain Division, there is a moderate positive correlation between SVI flags from 2010 and dollars spent on LIHTC in 2011-2020, with a Pearson’s score of 0.57.

To check to see if our correlations were not reflecting any major outliers, we used K-Means Clustering to separate our data points into meaningful groups. We then recalculated our Pearson's R for each group.

We utilized diff-in-diff models to analyze the impacts of the New Markets Tax Credit (NMTC) and the Low Income Housing Tax Credit (LIHTC) programs as they relate to social vulnerability and economic changes in Mountain Division tracts.Diff-In-Diff models are useful as a statistical tool to analyze the effects of a program when there is a treatment group and a group who did not receive the treatment (control group). It can be used when there are two periods, before intervention and after intervention. Since we have tracts who were eligible and did receive NMTC and LIHTC dollars and tracts who did not, we analyzed the impact of these programs before and after intervention.


# Results

## NMTC Diff-In-Diff Models

### Socioeconomic SVI

```{r, eval=FALSE}

write_model_outcomes(m1_nmtc_div, "NMTC", "socioeconomic status")

```
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_SES with treat, post and cbsa (formula: SVI_FLAG_COUNT_SES ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.18, F(75, 3360) = 10.16, p < .001, adj. R2 = 0.17)

The effect of treat × post is statistically non-significant and positive (beta = 0.02, 95% CI [-0.35, 0.39], t(3360) = 0.11, p = 0.914; Std. beta = 1.68e-03, 95% CI [-0.03, 0.03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on socioeconomic status-related social vulnerability and economic outcomes.


### Household Characteristics

```{r, eval=FALSE}
write_model_outcomes(m2_nmtc_div, "NMTC", "household characteristics")
```

Show in New Window
We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_HHCHAR with treat, post and cbsa (formula: SVI_FLAG_COUNT_HHCHAR ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.19, F(75, 3360) = 10.28, p < .001, adj. R2 = 0.17)

The effect of treat × post is statistically non-significant and positive (beta = 4.92e-03, 95% CI [-0.25, 0.26], t(3360) = 0.04, p = 0.970; Std. beta = 5.86e-04, 95% CI [-0.03, 0.03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on household characteristics-related social vulnerability and economic outcomes.


### Racial & Ethnic Minority SVI

```{r, eval=FALSE}
write_model_outcomes(m3_nmtc_div, "NMTC", "racial and ethnic minority status")
```

We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_REM with treat, post and cbsa (formula: SVI_FLAG_COUNT_REM ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.34, F(75, 3360) = 23.24, p < .001, adj. R2 = 0.33)

The effect of treat × post is statistically non-significant and negative (beta = -4.57e-03, 95% CI [-0.11, 0.10], t(3360) = -0.09, p = 0.932; Std. beta = -1.19e-03, 95% CI [-0.03, 0.03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on racial and ethnic minority status-related social vulnerability and economic outcomes.


### Housing & Transportation SVI

```{r, eval=FALSE}
write_model_outcomes(m4_nmtc_div, "NMTC", "housing and transportation access")
```

We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_HOUSETRANSPT with treat, post and cbsa (formula: SVI_FLAG_COUNT_HOUSETRANSPT ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and weak proportion of variance (R2 = 0.11, F(75, 3360) = 5.71, p < .001, adj. R2 = 0.09)

The effect of treat × post is statistically non-significant and positive (beta = 0.09, 95% CI [-0.19, 0.36], t(3360) = 0.62, p = 0.537; Std. beta = 0.01, 95% CI [-0.02, 0.04])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on housing and transportation access-related social vulnerability and economic outcomes.


### Overall SVI

```{r, eval=FALSE}
write_model_outcomes(m5_nmtc_div, "NMTC", "socioeconomic, household characteristics, racial and ethnic minority status, and housing and transportation access")
```

We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_OVERALL with treat, post and cbsa (formula: SVI_FLAG_COUNT_OVERALL ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.24, F(75, 3360) = 13.76, p < .001, adj. R2 = 0.22)

The effect of treat × post is statistically non-significant and positive (beta = 0.11, 95% CI [-0.64, 0.85], t(3360) = 0.28, p = 0.779; Std. beta = 4.23e-03, 95% CI [-0.03, 0.03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on socioeconomic, household characteristics, racial and ethnic minority status, and housing and transportation access-related social vulnerability and economic outcomes.


### Median Income Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m6_nmtc_div, "NMTC", "Median Income")
```

We fitted a linear model (estimated using OLS) to predict MEDIAN_INCOME with treat, post and cbsa (formula: MEDIAN_INCOME ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.17, F(75, 3360) = 9.26, p < .001, adj. R2 = 0.15)

The effect of treat × post is statistically non-significant and positive (beta = 0.03, 95% CI [-0.04, 0.10], t(3360) = 0.89, p = 0.373; Std. beta = 0.01, 95% CI [-0.02, 0.04])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on Median Income-related social vulnerability and economic outcomes.


### Median Home Value Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m7_nmtc_div, "NMTC", "Median Home Value")
```


We fitted a linear model (estimated using OLS) to predict MEDIAN_HOME_VALUE with treat, post and cbsa (formula: MEDIAN_HOME_VALUE ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.26, F(75, 3248) = 14.99, p < .001, adj. R2 = 0.24)

The effect of treat × post is statistically non-significant and positive (beta = 0.05, 95% CI [-0.07, 0.17], t(3248) = 0.88, p = 0.381; Std. beta = 0.01, 95% CI [-0.02, 0.04])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on Median Home Value-related social vulnerability and economic outcomes.


### House Price Index Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m8_nmtc_div, "NMTC", "House Price Index")
```

We fitted a linear model (estimated using OLS) to predict HOUSE_PRICE_INDEX with treat, post and cbsa (formula: HOUSE_PRICE_INDEX ~ treat + post + treat * post + cbsa) where treat represents NMTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.61, F(71, 1902) = 41.96, p < .001, adj. R2 = 0.60)

The effect of treat × post is statistically non-significant and negative (beta = -5.21e-04, 95% CI [-0.11, 0.11], t(1902) = -9.32e-03, p = 0.993; Std. beta = -1.33e-04, 95% CI [-0.03, 0.03])

Since the effect of treat x post is not statistically significant, we cannot conclude that the NMTC program had a measurable impact on House Price Index-related social vulnerability and economic outcomes.


## LIHTC Diff-In-Diff Models

### Socioeconomic SVI

```{r, eval=FALSE}
write_model_outcomes(m1_lihtc_div, "LIHTC", "socioeconomic status")
```

We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_SES with treat, post and cbsa (formula: SVI_FLAG_COUNT_SES ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.26, F(39, 332) = 2.91, p < .001, adj. R2 = 0.17)

The effect of treat × post is statistically non-significant and negative (beta = -0.26, 95% CI [-0.87, 0.35], t(332) = -0.83, p = 0.406; Std. beta = -0.04, 95% CI [-0.13, 0.05])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on socioeconomic status-related social vulnerability and economic outcomes.


### Household Characteristics SVI

```{r, eval=FALSE}
write_model_outcomes(m2_lihtc_div, "LIHTC", "household characteristics")
```


We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_HHCHAR with treat, post and cbsa (formula: SVI_FLAG_COUNT_HHCHAR ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.43, F(39, 332) = 6.52, p < .001, adj. R2 = 0.37)

The effect of treat × post is statistically non-significant and negative (beta = -0.06, 95% CI [-0.64, 0.53], t(332) = -0.19, p = 0.847; Std. beta = -7.97e-03, 95% CI [-0.09, 0.07])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on household characteristics-related social vulnerability and economic outcomes.

### Racial and Ethnic Minority SVI

```{r, eval=FALSE}
write_model_outcomes(m3_lihtc_div, "LIHTC", "racial and ethnic minority status")
```


We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_REM with treat, post and cbsa (formula: SVI_FLAG_COUNT_REM ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.59, F(39, 332) = 12.22, p < .001, adj. R2 = 0.54)

The effect of treat × post is statistically non-significant and negative (beta = -6.17e-03, 95% CI [-0.20, 0.19], t(332) = -0.06, p = 0.950; Std. beta = -2.23e-03, 95% CI [-0.07, 0.07])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on racial and ethnic minority status-related social vulnerability and economic outcomes.

### Housing and Transportation SVI

```{r, eval=FALSE}
write_model_outcomes(m4_lihtc_div, "LIHTC", "housing and transportation access")
```

We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_HOUSETRANSPT with treat, post and cbsa (formula: SVI_FLAG_COUNT_HOUSETRANSPT ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and moderate proportion of variance (R2 = 0.20, F(39, 332) = 2.08, p < .001, adj. R2 = 0.10)

The effect of treat × post is statistically non-significant and positive (beta = 0.19, 95% CI [-0.37, 0.75], t(332) = 0.67, p = 0.503; Std. beta = 0.03, 95% CI [-0.06, 0.13])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on housing and transportation access-related social vulnerability and economic outcomes.


### Overall SVI

```{r, eval=FALSE}
write_model_outcomes(m5_lihtc_div, "LIHTC", "socioeconomic, household characteristics, racial and ethnic minority status, and housing and transportation access")
```

We fitted a linear model (estimated using OLS) to predict SVI_FLAG_COUNT_OVERALL with treat, post and cbsa (formula: SVI_FLAG_COUNT_OVERALL ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.42, F(39, 332) = 6.20, p < .001, adj. R2 = 0.35)

The effect of treat × post is statistically non-significant and negative (beta = -0.13, 95% CI [-1.39, 1.12], t(332) = -0.21, p = 0.837; Std. beta = -8.60e-03, 95% CI [-0.09, 0.07])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on socioeconomic, household characteristics, racial and ethnic minority status, and housing and transportation access-related social vulnerability and economic outcomes.


### Median Income Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m6_lihtc_div, "LIHTC", "Median Income")
```

We fitted a linear model (estimated using OLS) to predict MEDIAN_INCOME with treat, post and cbsa (formula: MEDIAN_INCOME ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.32, F(39, 332) = 4.10, p < .001, adj. R2 = 0.25)

The effect of treat × post is statistically non-significant and positive (beta = 0.04, 95% CI [-0.20, 0.27], t(332) = 0.33, p = 0.745; Std. beta = 0.01, 95% CI [-0.07, 0.10])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on Median Income-related social vulnerability and economic outcomes.


### Median Home Value Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m7_lihtc_div, "LIHTC", "Median Home Value")
```

We fitted a linear model (estimated using OLS) to predict MEDIAN_HOME_VALUE with treat, post and cbsa (formula: MEDIAN_HOME_VALUE ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.48, F(39, 302) = 7.11, p < .001, adj. R2 = 0.41)

The effect of treat × post is statistically non-significant and positive (beta = 0.04, 95% CI [-0.25, 0.32], t(302) = 0.26, p = 0.798; Std. beta = 0.01, 95% CI [-0.07, 0.09])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on Median Home Value-related social vulnerability and economic outcomes.


### House Price Index Economic Outcomes

```{r, eval=FALSE}
write_model_outcomes(m8_lihtc_div, "LIHTC", "House Price Index")
```

We fitted a linear model (estimated using OLS) to predict HOUSE_PRICE_INDEX with treat, post and cbsa (formula: HOUSE_PRICE_INDEX ~ treat + post + treat * post + cbsa) where treat represents LIHTC program participation, post is the year of 2020 after starting period of 2010, and cbsa controls for metro-level effects.

The model explains a statistically significant and substantial proportion of variance (R2 = 0.65, F(23, 102) = 8.41, p < .001, adj. R2 = 0.58)

The effect of treat × post is statistically non-significant and positive (beta = 0.03, 95% CI [-0.37, 0.44], t(102) = 0.17, p = 0.865; Std. beta = 9.99e-03, 95% CI [-0.11, 0.13])

Since the effect of treat x post is not statistically significant, we cannot conclude that the LIHTC program had a measurable impact on House Price Index-related social vulnerability and economic outcomes.


# Discussion and Recommendations

In conclusion, for the Mountain Region, there was a a moderate to high correlation between tracts that received funding and those in need. However, there was no statistically significant impact of these programs on decreasing social vulnerability in any category.

Our analysis would indicate that the areas of need received funding. However, the amount of funding greatly varied across tracts. The amount of money should be included in our analysis to determine if the funding amount affected neighborhood vulnerability. Additionally, it may be beneficial to track gentrification, and check variables that measure housing affordability and population mobility.


# References

## R and Package Libraries

```{r, eval=FALSE}
# Load packages
library(renv)         # R environment management
library(devtools)     # R developer tools
library(here)         # relative filepaths for reproducibility
library(tidyverse)    # data wrangling
library(stringi)      # string wrangling
library(kableExtra)   # table formatting
library(tidycensus)   # census data
library(ggplot2)      # data visualization
library(patchwork)    # data visualization layout
library(magick)       # image editing in R
library(ggiraph)      # interactive data visualization
library(htmlwidgets)  # create widget from interactive visualizations
library(htmltools)    # encode HTML code in map tooltips
library(widgetframe)  # save widget of interactive visualizations
library(showtext)     # utilize special fonts in graphs
library(scales)       # format numbers and colors for graphs
library(tigris)       # pull US Census shapefiles
library(cowplot)      # data visualization plotting
library(gridExtra)    # grid for data visualizations
library(biscale)      # bivariate mapping
library(cluster)      # clustering algorithms
library(factoextra)   # clustering algorithms & visualization
library(rio)          # read excel file from URL
library(unhcrthemes)  # data visualization themes
library(ggrepel)      # data visualization formatting to avoid overlapping
library(rcompanion)   # data visualization of variable distribution
library(ggpubr)       # data visualization of variable distribution
library(moments)      # measures of skewness and kurtosis
library(tinytable)    # format regression tables
library(modelsummary) # format regression tables
library(report)    # package for statistical report writing and citations
```

```{r, eval=FALSE}
session <- sessionInfo()

report_system(session)
```

## R Version

Analyses were conducted using the R Statistical language (version 4.4.1; R Core Team, 2024) on Windows 11 x64 (build
26100)

```{r, eval=FALSE}
cite_packages(session, include_R = TRUE)
```

## R Packages

 - Arel-Bundock V (2022). “modelsummary: Data and Model Summaries in R.” _Journal of Statistical Software_, *103*(1), 1-23. doi:10.18637/jss.v103.i01 <https://doi.org/10.18637/jss.v103.i01>.
  - Arel-Bundock V (2025). _tinytable: Simple and Configurable Tables in 'HTML', 'LaTeX', 'Markdown', 'Word', 'PNG', 'PDF', and 'Typst' Formats_. R package version 0.8.0, <https://vincentarelbundock.github.io/tinytable/>.
  - Auguie B (2017). _gridExtra: Miscellaneous Functions for "Grid" Graphics_. R package version 2.3.
  - Chan C, Leeper T, Becker J, Schoch D (2023). _rio: A Swiss-army knife for data file I/O_. <https://cran.r-project.org/package=rio>.
  - Cheng J, Sievert C, Schloerke B, Chang W, Xie Y, Allen J (2024). _htmltools: Tools for HTML_. R package version 0.5.8.1, https://rstudio.github.io/htmltools/, <https://github.com/rstudio/htmltools>.
  - Gagolewski M (2022). “stringi: Fast and portable character string processing in R.” _Journal of Statistical Software_, *103*(2), 1-59. doi:10.18637/jss.v103.i02 <https://doi.org/10.18637/jss.v103.i02>.
  - Gohel D, Skintzos P (2025). _ggiraph: Make 'ggplot2' Graphics Interactive_. R package version 0.8.13, <https://davidgohel.github.io/ggiraph/>.
  - Grolemund G, Wickham H (2011). “Dates and Times Made Easy with lubridate.” _Journal of Statistical Software_, *40*(3), 1-25. <https://www.jstatsoft.org/v40/i03/>.
  - Karambelkar B (2017). _widgetframe: 'Htmlwidgets' in Responsive 'iframes'_. R package version 0.3.1, https://bhaskarvk.github.io/widgetframe/, <https://github.com/bhaskarvk/widgetframe>.
  - Kassambara A (2023). _ggpubr: 'ggplot2' Based Publication Ready Plots_. R package version 0.6.0, <https://rpkgs.datanovia.com/ggpubr/>.
  - Kassambara A, Mundt F (2020). _factoextra: Extract and Visualize the Results of Multivariate Data Analyses_. R package version 1.0.7, <http://www.sthda.com/english/rpkgs/factoextra>.
  - Komsta L, Novomestky F (2022). _moments: Moments, Cumulants, Skewness, Kurtosis and Related Tests_. R package version 0.14.1, http://www.komsta.net/, <https://www.r-project.org>.
  - Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2025). _cluster: Cluster Analysis Basics and Extensions_. R package version 2.1.8.1 - For new features, see the 'NEWS' and the 'Changelog' file in the package source), <https://CRAN.R-project.org/package=cluster>.
  - Makowski D, Lüdecke D, Patil I, Thériault R, Ben-Shachar M, Wiernik B (2023). “Automated Results Reporting as a Practical Tool to Improve Reproducibility and Methodological Best Practices Adoption.” _CRAN_. <https://easystats.github.io/report/>.
  - Mangiafico SS (2025). _rcompanion: Functions to Support Extension Education Program Evaluation_. Rutgers Cooperative Extension, New Brunswick, New Jersey. version 2.5.0, <https://CRAN.R-project.org/package=rcompanion/>.
  - Müller K (2020). _here: A Simpler Way to Find Your Files_. R package version 1.0.1, https://github.com/r-lib/here, <https://here.r-lib.org/>.
  - Müller K, Wickham H (2023). _tibble: Simple Data Frames_. R package version 3.2.1, https://github.com/tidyverse/tibble, <https://tibble.tidyverse.org/>.
  - Ooms J (2025). _magick: Advanced Graphics and Image-Processing in R_. R package version 2.8.6, <https://docs.ropensci.org/magick/https://ropensci.r-universe.dev/magick>.
  - Pedersen T (2024). _patchwork: The Composer of Plots_. R package version 1.3.0, https://github.com/thomasp85/patchwork, <https://patchwork.data-imaginist.com>.
  - Prener C, Grossenbacher T, Zehr A (2022). _biscale: Tools and Palettes for Bivariate Thematic Mapping_. R package version 1.0.0, <https://chris-prener.github.io/biscale/>.
  - Qiu Y, details. aotifSfAf (2020). _showtextdb: Font Files for the 'showtext' Package_. R package version 3.0.
  - Qiu Y, details. aotifSfAf (2024). _sysfonts: Loading Fonts into R_. R package version 0.8.9, <https://github.com/yixuan/sysfonts>.
  - Qiu Y, details. aotisSfAf (2024). _showtext: Using Fonts More Easily in R Graphs_. R package version 0.9-7, <https://github.com/yixuan/showtext>.
  - R Core Team (2024). _R: A Language and Environment for Statistical Computing_. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.
  - Slowikowski K (2024). _ggrepel: Automatically Position Non-Overlapping Text Labels with 'ggplot2'_. R package version 0.9.6, https://github.com/slowkow/ggrepel, <https://ggrepel.slowkow.com/>.
  - Ushey K, Wickham H (2025). _renv: Project Environments_. R package version 1.1.4, <https://CRAN.R-project.org/package=renv>.
  - Vaidyanathan R, Xie Y, Allaire J, Cheng J, Sievert C, Russell K (2023). _htmlwidgets: HTML Widgets for R_. R package version 1.6.4, <https://github.com/ramnathv/htmlwidgets>.
  - Vidonne C, Dicko A (2025). _unhcrthemes: UNHCR 'ggplot2' Theme and Colour Palettes_. R package version 0.6.3, https://unhcr-dataviz.github.io/unhcrthemes/, <https://github.com/unhcr-dataviz/unhcrthemes>.
  - Walker K (2025). _tigris: Load Census TIGER/Line Shapefiles_. R package version 2.2.0, commit 99b45062a24facbdbfdb6749ca531811d64a52b6, <https://github.com/walkerke/tigris>.
  - Walker K, Herman M (2025). _tidycensus: Load US Census Boundary and Attribute Data as 'tidyverse' and 'sf'-Ready Data Frames_. R package version 1.7.1, <https://walker-data.com/tidycensus/>.
  - Wickham H (2016). _ggplot2: Elegant Graphics for Data Analysis_. Springer-Verlag New York. ISBN 978-3-319-24277-4, <https://ggplot2.tidyverse.org>.
  - Wickham H (2023). _forcats: Tools for Working with Categorical Variables (Factors)_. R package version 1.0.0, https://github.com/tidyverse/forcats, <https://forcats.tidyverse.org/>.
  - Wickham H (2023). _stringr: Simple, Consistent Wrappers for Common String Operations_. R package version 1.5.1, https://github.com/tidyverse/stringr, <https://stringr.tidyverse.org>.
  - Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). “Welcome to the tidyverse.” _Journal of Open Source Software_, *4*(43), 1686. doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.
  - Wickham H, Bryan J, Barrett M, Teucher A (2024). _usethis: Automate Package and Project Setup_. R package version 3.1.0, https://github.com/r-lib/usethis, <https://usethis.r-lib.org>.
  - Wickham H, François R, Henry L, Müller K, Vaughan D (2023). _dplyr: A Grammar of Data Manipulation_. R package version 1.1.4, https://github.com/tidyverse/dplyr, <https://dplyr.tidyverse.org>.
  - Wickham H, Henry L (2025). _purrr: Functional Programming Tools_. R package version 1.0.4, https://github.com/tidyverse/purrr, <https://purrr.tidyverse.org/>.
  - Wickham H, Hester J, Bryan J (2024). _readr: Read Rectangular Text Data_. R package version 2.1.5, https://github.com/tidyverse/readr, <https://readr.tidyverse.org>.
  - Wickham H, Hester J, Chang W, Bryan J (2022). _devtools: Tools to Make Developing R Packages Easier_. R package version 2.4.5, https://github.com/r-lib/devtools, <https://devtools.r-lib.org/>.
  - Wickham H, Pedersen T, Seidel D (2023). _scales: Scale Functions for Visualization_. R package version 1.3.0, https://github.com/r-lib/scales, <https://scales.r-lib.org>.
  - Wickham H, Vaughan D, Girlich M (2024). _tidyr: Tidy Messy Data_. R package version 1.3.1, https://github.com/tidyverse/tidyr, <https://tidyr.tidyverse.org>.
  - Wilke C (2024). _cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2'_. R package version 1.1.3, <https://wilkelab.org/cowplot/>.
  - Zhu H (2024). _kableExtra: Construct Complex Table with 'kable' and Pipe Syntax_. R package version 1.4.0, https://github.com/haozhu233/kableExtra, <http://haozhu233.github.io/kableExtra/>.

## Data

-  CDFI Fund (2023). _FY 2023 NMTC Public Data Release: 2003-2021 Data File Updated - Aug 21, 2023_. https://www.cdfifund.gov/documents/data-releases

-  Centers for Disease Control and Prevention/ Agency for Toxic Substances and Disease Registry/ Geospatial Research, Analysis, and Services Program. (2022). _CDC/ATSDR Social Vulnerability Index 2020 Methodology_. https://web.archive.org/web/20241028180954/https://www.atsdr.cdc.gov/placeandhealth/svi/documentation/SVI_documentation_2020.html

-  FHFA (n.d.). _HPI® Census Tracts (Developmental Index; Not Seasonally Adjusted)_. https://www.fhfa.gov/DataTools/Downloads/Pages/House-Price-Index-Datasets.aspx#atvol

-  HUD User (n.d.). _2010, 2011, and 2012 QCT data for all of the census tracts in the United States and Puerto Rico (qct_data_2010_2011_2012.xlsx)_. https://www.huduser.gov/portal/datasets/qct.html#year2010

-  HUD User (2023). _Low-Income Housing Tax Credit (LIHTC): Property Level Data_. https://www.huduser.gov/portal/datasets/lihtc/property.html

-  Novogradac New Markets Tax Credit Resource Center. (2017). _New Markets Tax Credit Low-Income Community Census Tracts - American Community Survey 2011-2015_. https://www.novoco.com/resource-centers/new-markets-tax-credits/data-tables

-  Steven Manson, Jonathan Schroeder, David Van Riper, Katherine Knowles, Tracy Kugler, Finn Roberts, and Steven Ruggles. _IPUMS National Historical Geographic Information System: Version 18.0 [2020 → 2010 Block Groups → Census Tracts Crosswalks National File]_. Minneapolis, MN: IPUMS. 2023. http://doi.org/10.18128/D050.V18.0

-  U.S. Bureau of Labor Statistics (n.d.). _CPI Inflation Calculator_. https://data.bls.gov/cgi-bin/cpicalc.pl

-  U.S. Bureau of Labor Statistics (n.d.). _QCEW County-MSA-CSA Crosswalk (For NAICS-Based Data)_. https://www.bls.gov/cew/classifications/areas/county-msa-csa-crosswalk.htm

-  U.S. Census Bureau. (2011). _2006-2010 American Community Survey 5-year_. https://www.census.gov/newsroom/releases/archives/american_community_survey_acs/cb11-208.html

-  U.S. Census Bureau. (2013). _2008-2012 American Community Survey 5-year_.  https://www.census.gov/newsroom/press-kits/2013/20131217_acs_5yr.html

-  U.S. Census Bureau. (2022). _2016-2020 American Community Survey 5-year_. https://www.census.gov/newsroom/press-releases/2022/acs-5-year-estimates.html

## Readings

-  Jayachandran, A. (2024). What Is House Price Index (HPI)? https://www.wallstreetmojo.com/house-price-index-hpi/

-  Killingsworth, M. A. (2021). Experienced well-being rises with income, even above $75,000 per year. Proceedings of the National Academy of Sciences - PNAS, 118(4). https://doi.org/10.1073/pnas.2016976118_download

-  Tax Foundation (2020). An Overview of the Low-Income Housing Tax Credit (LIHTC) https://taxfoundation.org/research/all/federal/low-income-housing-tax-credit-lihtc/

-  The Urban Institute (2021). The Past, Present, and Future of the New Markets Tax Credit Program https://www.urban.org/events/past-present-and-future-new-markets-tax-credit-program


